---
title: "PRRSV-GSVA"
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,eval = FALSE)
```

## Summary

PRRSV is a porcine virus. The idea of the project is to use single sample gene set enrichment scores to correlate the early innate response with antibody response at a later timepoint. 24 animal animals were separated in 4 different groups:


Loading the necessary packages


```{r}
## First specify the packages of interest
packages = c("tidyverse","Biobase","limma","GSEABase","sigPathway","readxl","ggpubr")

## Now load or install&load all
package.check <- lapply(
  packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      BiocManager::install(x)
      library(x, character.only = TRUE)
    }
  }
)

```
## 
we can change the value for the day

day <- 7

```{r,echo=TRUE}
countfile<-read.delim(file = paste0("p470/Day",day,"/extended/p470.counts.txt"),check.names = FALSE,row.names = 1)
##the count file that is read by read.delim depends on the day argument above
```
small loop to create a genesets list
```{r,echo=FALSE}
##the countfile must be read as a matrix to be used in gsva function
eset_matrix <- as.matrix(countfile)
for (i in 1:346){
        
        genesets[i]<-btmgeneset[[i]]$title
}
##we create the btmgeneset using the gmx file
btmgeneset<-gmxToG("BTM gene sets.gmx")# from the sigpathway package
```

### Main function of the GSVA that will generate the Enrichment Scores
```{r,echo=TRUE}
results_gsva<-lapply(X = btmgeneset,FUN = function(x){gsva(expr=eset_matrix, 
     gset.idx.list=x, 
     annotation,
     method="gsva",
     kcdf="Poisson",#poisson for discrete numbers
     abs.ranking=FALSE,
     min.sz=1,#minimal size of each genesets
     max.sz=Inf,#maximal size of each genesets
     parallel.sz=0,
     parallel.type="SOCK",
     mx.diff=TRUE,
     tau=1,#tau=switch(method,gsva=1, ssgsea=0.25, NA),
     ssgsea.norm=TRUE,
     verbose=TRUE)})


```

```{r,echo=TRUE}
results_gsva_df<-data.frame(matrix(unlist(results_gsva),#we transform the list in a dataframe
                                   nrow = length(results_gsva),
                                   byrow = T),
                            stringsAsFactors = FALSE)

results_gsva_df<-t(results_gsva_df)#we transpose because we want the genesets as col and samples as rows

sampleid<-colnames(countfile)#we take the names of the samples

colnames(results_gsva_df)<- genesets

#add the rownames of the samples (now as rows)


##assign a new name to the data frame with Day + day of the count file + _ES (enrichment score)
```

```{r,echo=FALSE}
results_gsva_df$sampleid<-sampleid
assign(paste("Day",day,"_ES",sep = ""),results_gsva_df)
```

We bind the 3 data frames togther
```{r}
#we bind the 3 data frames by rows
completeES<-rbind(Day0_ES_t,Day3_ES_t,Day7_ES_t)
head(completeES)
dim(completeES)
```

```{r}
#we do a pivot of the Enrichment score (ES) to get only one column with our ES
completeES_long<-pivot_longer(completeES,
                              cols = -sampleid,
                              names_to = "genesets",
                              values_to = "es")

# we clean the sample id column
completeES_long$sampleid<-gsub("_","",completeES_long$sampleid)# we replace the "_" by ""
## we replace "CH431" by "" in order to use the ID number for merging the data set
completeES_long$sampleid<-gsub("CH431","",completeES_long$sampleid)
samplesinfos$sampleid<-gsub("CH431-","",samplesinfos$sampleid)
## we set the sampleid col to numeric in the ES and the sampleinfos file
completeES_long$sampleid<-as.numeric(completeES_long$sampleid)
```

```{r}
#import the data frame containing the info about the samples
samplesinfos <- read_excel("CH431.xlsx", 
                           col_names = FALSE, skip = 2,col_types = c("text"))
#we assign new names for the columns
colnames(samplesinfos)<- c("sampleid",
                           "virus",
                           "day",
                           "code",
                           "animalid",
                           "treatment",
                           "rqn")

samplesinfos$sampleid<-as.numeric(samplesinfos$sampleid)

#######Major step we merge the ES data with the sample informations ####

completeES_long_merge<-merge(completeES_long,samplesinfos)
```

We now have a dataframe with the infos of the samples with the enrichmentscore values
We proceed with cleaning of the data 

```{r}

completeES_long_merge$day<-gsub("Day","",completeES_long_merge$day)

#we remove the trailing whitespace
completeES_long_merge$day<-str_trim(completeES_long_merge$day)

#we transform to factors
completeES_long_merge$genesets<-as.factor(as.character(completeES_long_merge$genesets))
str(completeES_long_merge)
```

we take the proliferation data sets and do some cleaning
```{r}

proliferation <- read_excel("T cell data/Kick_PRRSV_proliferation.xlsx", 
                            skip = 1)
proliferation_longer<-pivot_longer(proliferation,cols=-1:-5)

colnames(proliferation_longer)<-tolower(colnames(proliferation_longer))

proliferation_longer$name<-as.factor(tolower(proliferation_longer$name))

names(proliferation_longer)<-c("well","day","treatment","stimulus","animalid","name","value" )

proliferation_longer$animalid<-factor(proliferation_longer$animalid)
proliferation_longer$day<-factor(as.character(proliferation_longer$day))

str(proliferation_longer)
head(proliferation_longer)
```

```{r}
#we clean certain columns 

completeES_long_merge$treatment<-word(completeES_long_merge$treatment,1)#we extract the treatment value in a simpler form now ("MLV","MOCK"...)
completeES_long_merge$day<-factor(as.character(completeES_long_merge$day))
completeES_long_merge$animalid<-factor(completeES_long_merge$animalid)


#for the last step we merge together the data sets ES and proliferation data in 1 super data frame called superdf

superdf<-merge(proliferation_longer,completeES_long_merge,by=c("animalid","treatment"),all = TRUE)
head(superdf)
str(superdf)
##create a superdf using the ratio between the days7 and day3 /day0
## we proliferation_clean which contains only the homologous treatment MOCK MOCK HP HP etc
superdf_ratio_clean<-merge(proliferation_longer_clean,completeES_long_merge_ratio,by=c("animalid"),all = TRUE)
```

## Including Plots

You can also embed plots, for example:

```{r , echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
